Piecewise Regression for Nutritional Outcomes Pre- and Post-CFTR Modulator

### 2022-05-15 StatNet/knot_models.qmd
Update knot_models.qmd
bayesian approach on triceps data

---
title: "Piecewise Regression for Nutritional Outcomes Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
format: revealjs
editor: visual
bibliography: /Users/timvigers/Dropbox/Miscellaneous/references.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association-alphabetical.csl
---

```{r}
#| include: false
# Load libraries, data, etc.
library(DiagrammeR)
library(MultiKink)
library(ggplot2)
library(broom)
library(knitr)
data("triceps")
cf = read.csv("/Users/timvigers/GitHub/CF/StatNet/data.csv",
              stringsAsFactors = F,na.strings = "")
```

@@

```{r}
#| eval: false
grViz("
g [label = '@@7']
h [label = '@@8']
i [label = '@@9']
j [label = '@@10']
k [label = '@@11']
l [label = '@@12']
m [label = '@@13']

a -> b
b -> c
c -> d
c -> e
c -> f
d -> g
e -> h
f -> i
g -> j
h -> k
i -> l
j -> m
k -> m
l -> m
}

[1]: paste0('Clinical Database (n = ', ncdb, ')')
[2]: paste0('Treated With Ivacaftor or Trikafta From 2008-2021 (n = ',length(unique(mods$`Patient ID`)), ')')
[3]: paste0('In Our Dataset (n = ',length(intersect(mods$`Patient ID`,sweat$`Patient ID`)), ')')
[4]: paste0('Treated With Ivacaftor Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Kalydeco']), ')')
[5]: paste0('Treated With Trikafta Alone From 2008-2021 (n = ',as.numeric(table(groups$Group)['Trikafta']), ')')
[6]: paste0('Treated With Ivacaftor followed by Trikafta From 2008-2021 (n = ',as.numeric(table(groups$Group)['Both']), ')')
[7]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Kalydeco']), ')')
[8]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Trikafta']), ')')
[9]: paste0('At Least One Vitamin A Measures Prior to First Modulator (n = ',as.numeric(table(two_measures$Group)['Both']), ')')
[10]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Kalydeco']), ')')
[11]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Trikafta']), ')')
[12]: paste0('At Least One Vitamin A >= 3 Months After First Modulator (n = ',as.numeric(table(after_measure$Group)['Both']), ')')
[13]: paste0('Final (n = ',length(unique(after_measure$`Patient ID`)), ')')
")
```

@@

>>>>

---
title: "Piecewise Regression for Nutritional Outcomes Pre- and Post-CFTR Modulator"
author: "Tim Vigers"
format: revealjs
editor: visual
bibliography: /Users/timvigers/Dropbox/Miscellaneous/references.bib
csl: /Users/timvigers/GitHub/styles/american-medical-association-alphabetical.csl
cache: false
---

```{r}
#| include: false
# Load libraries, data, etc.
library(DiagrammeR)
library(MultiKink)
library(ggplot2)
library(broom)
library(lmerTest)
library(mcp)
library(knitr)
data("triceps")
cf = read.csv("/Users/timvigers/GitHub/CF/StatNet/data.csv",
              stringsAsFactors = F,na.strings = "")
```

@@

-   Limited to 3 years pre/post

@@

-   The `mcp` package provides an easy interface for fitting Bayesian regression models with multiple change points.

## On the triceps data

```{r}
# Define the model
model = list(
  triceps ~ 1 + age,          # intercept + slope
  ~ 0 + age,  # joined slope with 3 change points
  ~ 0 + age,
  ~ 0 + age
)
bayes_fit = mcp(model, data = triceps, sample = "both")
plot(bayes_fit)
```

-   Visually, this approach seems to generally confirm the knots we used earlier.

-   The model results bear this out as well:

```{r}
s = summary(bayes_fit)
kable(s,digits = 3)
```

---

-   `mcp` also allows us to test hypotheses about the location of these knots:

```{r}
hypothesis(bayes_fit, "cp_1 = 5")
hypothesis(bayes_fit, "cp_2 = 10")
hypothesis(bayes_fit, "cp_3 = 20")
```

- None of the hypothesis tests are significant, indicating that the knots found with `mcp` are reasonably close to the ones we pre-specified (the CI on cp_1 is fairly wide though).

---

```{r}
# Define the model
model = list(
  Vitamin.A ~ 1 + Time, # intercept + slope
  ~ 0 + Time  # joined slope
)
fit = mcp(model, data = cf)
plot(fit)
```

-   What does this mean? Why is the change point estimated to be before 0?

---

```{r}
# Dirichlet prior on change point
d_prior = list(
  cp_1 = "dnorm(0,1)"
)
fit_cp_prior = mcp(model, prior = d_prior,data = cf)
plot(fit_cp_prior)
```

---
